Combining neural networks and geostatistics for landslide hazard assessment of San Salvador Metropolitan Area, El Salvador \newline \newline Combinando redes neuronales y geoestad\'{i}stica para evaluaci\'{o}n de deslizamientos de tierra del \'{a}rea Metropolitana de San Salvador, El Salvador

### 2015-11-17 paper.tex

All rocks exposed in MASS have volcanic origin and consist of intercaled primary and reworked deposits from Late Tertiary to Holocene age \cite{schmidt1975}. The younger ones are known as Tierra Blanca pyroclastic ash deposits and form a relative continuous layer in the majority of MASS with an average thickness of 4 m \cite{schmidt1975}. These are deposits derived mainly from Ilopango caldera eruptions with more than 50 m of thick \cite{schmidt1975}. Tierra Blanca soils are highly susceptible to erosion. One of the most common dangerous effects of the erosion in this area are landslides during heavy rainfall and earthquake ground shaking \cite{schmidt1975}. Ch\'{a}vez et al. \cite{chavez2014a} identified high and moderate erosion hazard in the majority of MASS area. The more recent layer Tierra Blanca, known as Tierra Blanca Joven (TBJ) \cite{hernan2004} presents very poor geotechnical properties, especially when there is an increment of soil moisture (i.e. during rainy seasons). This results in high instability and susceptibility (\cite{chavez2014b} and \cite{rolo2004}). Several incised rivers and ravines cross MASS area. Due to the large erosion rates of the area, natural slopes of these watercourses are often close to vertical and can reach heights of over 10 m. Taking into account this property of volcanic soils, usually vertical slopes are cut to urban and road construction. As Bommer \& Rodriguez \cite{bommer} identified in Central America, although such slopes may remain stable for years, they may become unstable, abruptly and totally under the action of heavy rainfall or seismic shaking. Hern\'{a}ndez \cite{hernan2004} carried out a detailed description of this process in TBJ, where these instabilities are very common. 
In the study area fault trends are characterized by steep angles (65\textdegree - 90\textdegree). The main fault system is characterized by an east-west trend; this trend is responsible for the steep northern slope of the El Balsamo Range.  The secondary systems are with north, northwest and northeast trends offset the main fault system and are more active in the present tectonic setting \cite{schmidt1975}. However, all the mentioned fault families appear active. Thus the several fault systems were formed at different times, but were repeatedly reactivated \cite{rymer1987}. Finally there are also several ringlike faults that can be related to subsurface collapse due to volcanic tectonic subsidence \cite{schmidt1975}. The last eruption of San Salvador volcano was in 1917.  The majority of the magma's total volume (97\%) extruded from several vents in the northwest flank of the volcano, outside of MASS. However, a fraction of MASS is located under the proximal volcanic hazard zone according to Sofield \cite{sofield2004}. In the steep slopes of San Salvador volcano, especially at Picacho, the highest peak which is located in the eastern side, there is a high lahar hazard that threatens MASS \cite{major2004}. In addition to that one of the most destructive eruptions in the world history was 430 A.D. The TBJ eruption was produced by the Ilopango Caldera and  according to Dull \cite{dull2004}, yielded total devastation in whole MASS area.

>>>>

All rocks exposed in MASS have volcanic origin and consist of intercaled primary and reworked deposits from Late Tertiary to Holocene age \cite{schmidt1975}. The younger ones are known as Tierra Blanca pyroclastic ash deposits and form a relative continuous layer in the majority of MASS with an average thickness of 4 m \cite{schmidt1975}. These are deposits derived mainly from Ilopango caldera eruptions with more than 50 m of thick \cite{schmidt1975}. Tierra Blanca soils are highly susceptible to erosion. One of the most common dangerous effects of the erosion in this area are landslides during heavy rainfall and earthquake ground shaking \cite{schmidt1975}. Ch\'{a}vez et al. \cite{chavez2014a} identified high and moderate erosion hazard in the majority of MASS area. The more recent layer Tierra Blanca, known as Tierra Blanca Joven (TBJ) \cite{hernan2004} presents very poor geotechnical properties, especially when there is an increment of soil moisture (i.e. during rainy seasons). This results in high instability and susceptibility (\cite{chavez2014b} and \cite{rolo2004}). Several incised rivers and ravines cross MASS area. Due to the large erosion rates of the area, natural slopes of these watercourses are often close to vertical and can reach heights of over 10 m. Taking into account this property of volcanic soils, usually vertical slopes are cut to urban and road construction. As Bommer \& Rodriguez \cite{bommer} identified in Central America, although such slopes may remain stable for years, they may become unstable, abruptly and totally under the action of heavy rainfall or seismic shaking. Hern\'{a}ndez \cite{hernan2004} carried out a detailed description of this process in TBJ, where these instabilities are very common. 
In the study area fault trends are characterized by steep angles (65\textdegree - 90\textdegree). The main fault system is characterized by an east-west trend; this trend is responsible for the steep northern slope of the El Balsamo Range. The other fault systems have north, northwest, and northeast trends and are more active in the present tectonic setting \cite{schmidt1975}. However, all the mentioned fault families appear active. Thus the several fault systems were formed at different times, but were repeatedly reactivated \cite{rymer1987}. Finally there are also several ringlike faults that can be related to subsurface collapse due to volcanic tectonic subsidence \cite{schmidt1975}. The last eruption of San Salvador volcano was in 1917.  The majority of the magma's total volume (97\%) extruded from several vents in the northwest flank of the volcano, outside of MASS. However, a fraction of MASS is located under the proximal volcanic hazard zone according to Sofield \cite{sofield2004}. In the steep slopes of San Salvador volcano, especially at Picacho, the highest peak which is located in the eastern side, there is a high lahar hazard that threatens MASS \cite{major2004}. In addition to that one of the most destructive eruptions in the world history was 430 A.D. The TBJ eruption was produced by the Ilopango Caldera and  according to Dull \cite{dull2004}, yielded total devastation in whole MASS area.
